---
title: "Profit"
author: "Amala Ann Mathew"
date: "2024-06-09"
output: word_document
---

```{r}
#Importing the necessary libraries:
library(tseries)
library(astsa)
library(forecast)
library(seastests)
```


```{r}
#Reading the data:
data <- read.csv(file.choose())
head(data)
attach(data)
```


```{r}
#Shapiro-Wilk test for normality:

#H0: Data is normal
#H1: Data is not normal

shapiro.test(Profit) 
```
Interpretation:
The Shapiro-Wilk test is a hypothesis test that is applied to a sample with a null hypothesis that the sample has been generated from a normal distribution. If the P-value is low(<0.05),we can reject the null hypothesis and say that the sample has not been generated from a normal distribution.

Since the P- value is greater than 0.05, the observations of the Sales column are normally distributed as indicated by the Shapiro-Wilk test [p-value is 0.1065].

```{r}
#Since the given data is a data frame we need to convert it into a time series object for an effective time series analysis 
df <- ts(Profit, start = 2018,  frequency = 12)
df
class(df)
```
Interpretation:
The data frame is now transformed into Time series data

```{r}
# Time series plot
ts.plot(df)
```
Interpretation:
From the time series plot it is observed that there is  no seasonal component.

```{r}
# Checking for seasonality:
isSeasonal(df, test = "combined",freq = 12)
isSeasonal(df, test = "qs", freq = 12)
isSeasonal(df, test = "kw", freq = 12)
isSeasonal(df, test = "fried", freq = 12)
```
Interpretation:
Mathematically proved that there doesn't exist seasonality. 

```{r}
# Autocorrelation function plot:
acf(df, main="ACF Plot of the Profit data")
```
Interpretation:
From the ACF plot it is observed that there is no significant lags. That means Non-stationary.
```{r}
#Augmented Dickey-Fuller Test

#HO:Data is non-stationary
#H1:Data is stationary
adf.test(df)
```
Interpretation:
Since the p-value is less than 0.05 we reject the null hypothesis. This means that the data is stationary.
```{r}
#stationary model
#finding the best ARIMA model 
fit = auto.arima(df , seasonal = FALSE) 
fit
```
Interpretation:
The ARIMA(2,1,0)  with the specified coefficients and diagnostic values suggests a reasonable fit for forecasting future values.

```{r}
# Forecasting can only be done only if residual analysis is satisfied:

#Residual analysis for model 1:
r = resid(fit)

acf(r, main = "ACF plot of the residuals")            


```
Interpretation:
The ACF of the residuals are non-significant at different lags,that is the residuals do not exhibit any systematic patterns or correlation. Therefore the the model has adequately captured the underlying structure of the time series data. 

```{r}
#Shapiro-Wilk test to check the normality of the residuals:
#H0: Residuals are normal.
#H1: Residuals are not normal.
shapiro.test(r) 

#Q-Q plot:
qqnorm(r)
qqline(r)

```
Interpretation:
Since the p-value is greater than 0.05 we fail to reject the null hypothesis.Therefore the residual is normal

```{r}
#Box-Pierce test to examine the null hypothesis of independence.
#H0: There is no significant autocorrelation.
#H1: There is some significant autocorrelation.
Box.test(r) #lag = p + q; p=1, q= 1       
#p-value > 0.05 -> fail to reject null hypothesis -> uncorrelated
```
Since the value is greater than 0.05 we fail to reject the null hypothesis. Therefore the residuals are uncorrelated
```{r}
# Forecasting:
forecast(fit, h = 4)    #h: no. of observations to be forecast.
```

The forecasted values of profit during first quarter of 2022 are given above.
